Low-Cost Data Deduplication for Virtual Machine Backup in Cloud Storage

### 2013-03-10 hotstorage13/arch.tex
revision

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level duplicate detection requires a global comparison of 
content fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a pre-scanning of  data  blocks in dirty VM segments  and after that, real backup  for those non-duplicates
is conducted. That does incur an extra  round of local VM block reading while avoiding inline 
deduplication and leading to a much smaller resource requirement.  
During VM reading, detection requests can be accumulated on the disk 
for a period time  in a  partitioned manner.  Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, the cost to 
process detection requests within a partition is significantly smaller than processing all the requests globally.

\item {\bf Buffered data redistribution in parallel duplicate detection}.  
Let {\em global index} be the meta data containing the fingerprint values of unique blocks of snapshots
in all VMs and  the reference pointers which lead to the location of raw data.
%A duplicate detection request for a content block needs to be compared with the global index to determine if this
%block is non-duplicate. If it is a duplicate, return the corresponding reference pointer.
%We conduct parallel processing of detection requests in all machines involved. 
A logical way to distribute detection requests among machines is to partition based on fingerprint values of content blocks
in a  disjointed manner.
Initial VM data reading follows the VM distribution among machines and the detected duplicate summary results 
need to be collected following the VM distribution. 
Therefore, there are two all-to-all data redistribution operations involved.
One is to map detection requests from VM-based distribution to fingerprint based distribution.  
One is to map duplicate summary from fingerprint-based distribution to VM based distribution.  
The redistributed data needs to be accumulated on the disk to minimize the use of memory.
To minimize the excessive disk seek cost, outgoing or incoming data exchange messages need to be buffered to bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, space per each buffer  is small under the memory constraint for large $p$ or $q$ values.
This counteracts the effort of seek cost reduction.  
%Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an efficient data exchange and disk data buffering  scheme to address this.

@@

\begin{figure*}[tbhp]
\centering
\includegraphics[width=0.75\textwidth]{steps.pdf}
\caption{Processing flow of Stages  1,  2 and 3.}
\label{fig:flow}
\end{figure*}

%\subsection{Processing flow and resource usage}
The data flow of our multi-stage duplicate detection is depicted in Figure~\ref{fig:flow}. 
%let $v$ be the total number of virtual machines hosted in  a cluster,  
%we divide the backup into $k$  iterations. 
%\begin{itemize}
%\item
In Stage 1, each machine independently reads  
%$v/k$  
VM images that need a backup
and forms duplicate  detection requests. For those segments that have been modified since last backup indicated by 
dirty bits,  the system divides  each segment into a sequence of chunk blocks,  computes the meta 
information such as chunk fingerprints,  sends a request to a proper machine, accumulates  
received requests into a partition on the local temporary disk storage. 
The partition mapping is using a hash function applied to the content fingerprint. 
Assuming all machines have a  homogeneous resource configuration, each machine is evenly  assigned with
$q$ parititons of content hash and accumulates corresponding requests on the disk. 
There are two options to assign buffers in each machine to buffer requests for each partition.
1) Each machine has  $p\times q$ send buffers corresponding to $p\times q$ partitions in the cluster
since a content block in a VM image of this machine can be sent to any of these partitions.
2) Each machine allocates $p$ send buffers to deliver requests to $p$ machines and 
$p$ receive buffers to collect requests  from other machines.
Then the system copies requests from each of $p$ receive buffers to  $q$ local request buffers,
and outputs each request buffer to one of the request partitions on the disk
when this request buffer becomes full.  Option 2 is much more efficient than Option 1 because $2p+q$ is much smaller than
$p\times q$. As a result, each buffer has a bigger size to accumulate requests and that means
less disk seek overhead.

@@

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and output non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update, the system  may find some  blocks with the same fingerprint may be 
created redundantly. For example, two different VM blocks have the same content block fingerprint while the  global 
index does not have such a fingerprint. The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

>>>>

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level duplicate detection requires a global comparison of 
content fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a pre-scanning of  data  blocks in dirty VM segments  and after that, real backup  for those non-duplicates
is conducted. That does incur an extra  round of local VM block reading while avoiding inline 
deduplication and leading to a much smaller resource requirement.  
During VM reading, detection requests are accumulated on the disk 
for a period time  in a  partitioned manner.  Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, the cost to 
process detection requests within a partition is significantly smaller than processing all the requests globally.

\item {\bf Buffered data redistribution in parallel duplicate detection}.  
Let {\em global index} be the meta data containing the fingerprint values of unique blocks of snapshots
in all VMs and  the reference pointers which lead to the location of raw data.
%A duplicate detection request for a content block needs to be compared with the global index to determine if this
%block is non-duplicate. If it is a duplicate, return the corresponding reference pointer.
%We conduct parallel processing of detection requests in all machines involved. 
A logical way to distribute detection requests among machines is to partition based on fingerprint values of content blocks
in a  disjointed manner.
Initial VM data reading follows the VM distribution among machines and the detected duplicate summary results 
need to be collected following the VM distribution. 
Therefore, there are two all-to-all data redistribution operations involved.
One is to map detection requests from VM-based distribution to fingerprint based distribution.  
Another one  is to map duplicate summary from fingerprint-based distribution to VM based distribution.  
The redistributed data needs to be accumulated on the disk to minimize the use of memory.
To minimize the excessive disk seek cost, outgoing or incoming data exchange messages need to be buffered to bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, space per each buffer  is small under the memory constraint for large $p$ or $q$ values.
This counteracts the effort of seek cost reduction.  
%Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an efficient data exchange and disk data buffering  scheme to address this.

@@

\begin{figure*}[tbhp]
\centering

\includegraphics[width=0.95\textwidth]{steps.pdf}
\caption{Processing flow of Stage  1 (dirty segment scan and request accumulation), Stage 2 
(fingerprint comparison and summary output),  and Stage 3 (non-duplicate block backup).}
\label{fig:flow}
\end{figure*}

%\subsection{Processing flow and resource usage}
The data flow of our multi-stage duplicate detection is depicted in Figure~\ref{fig:flow}. 
%let $v$ be the total number of virtual machines hosted in  a cluster,  
%we divide the backup into $k$  iterations. 
%\begin{itemize}
%\item
In Stage 1, each machine independently reads  
%$v/k$  
VM images that need a backup
and forms duplicate  detection requests. For those segments that have been modified since last backup indicated by 
dirty bits,  the system divides  each segment into a sequence of chunk blocks,  computes the meta 
information such as chunk fingerprints,  sends a request to a proper machine, accumulates  
received requests into a partition on the local temporary disk storage. 
The partition mapping is using a hash function applied to the content fingerprint. 
Assuming all machines have a  homogeneous resource configuration, each machine is evenly  assigned with
$q$ parititons of global index and accumulates corresponding requests on the disk. 
There are two options to allocate buffers at each machine. 
1) Each machine has  $p\times q$ send buffers corresponding to $p\times q$ partitions in the cluster
since a content block in a VM image of this machine can be sent to any of these partitions.
2) Each machine allocates $p$ send buffers to deliver requests to $p$ machines; it allocates 
$p$ receive buffers to collect requests  from other machines.
Then the system copies requests from each of $p$ receive buffers to  $q$ local request buffers,
and outputs each request buffer to one of the request partitions on the disk
when this request buffer becomes full.  Option 2 depicted in Figure~\ref{fig:flow}
is much more efficient than Option 1 because $2p+q$ is much smaller than
$p\times q$. As a result, each buffer in Option 2 has a bigger size to accumulate requests and that means
less disk seek overhead.

@@

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and outputs non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update after new block creation in the backup storage,
the system  may find some  blocks with the same fingerprint may be 
created redundantly. For example, two different VM blocks that have the same  fingerprint are not detected
because  the  global index has not contained such a fingerprint yet. 
The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

### 2013-03-10 hotstorage13/eval.tex
revision

We have implemented and evaluated a prototype of our multi-stage deduplication scheme on a Linux cluster
of multi-core AMD Bulldozer FX8120 and Intel Nehalem E5530 machines.  Objectives of our experimental evaluation are:
1) Analyze the deduplication throughput and effectiveness for a large number of VMs.
%Compare with the data domain approach~\cite{bottleneck08}.
2) Examine the impacts of buffering during metadata exchange.

@@

Our implementation is based on Alibaba's Xen cloud platform~\cite{Aliyun,WeiZhangIEEE}
where each multi-core machine   hosts  up to 25 VMs. 
%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.
The backup of VM snapshots is completed within a few  hours every night.
Based on our study of its production  data,  each VM has about  40GB of storage  data usage on average
including OS and user data disk. 
%and that translates to an aggregated backup throughput of 139GB per second, or 500TB per hour.
%  For 2 hours, every machine can do 139MB/second. For 3 hours, then 92MB/second

@@

We have performed a trace-driven study using  a 1323 VM dataset  collected from 100 Aliyun's cloud nodes~\cite{WeiZhangIEEE}.
%In this dataset, there are 10 snapshots per each VM user and the total amount of space 
%investigated is 17.5 terabytes.
%This dataset contas compose of 35 VMs from 7 popular OSes: 
%Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. For each OS, 
%5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
%The overall data size for this 700 full snapshots is 17.6 TB.
%This dataset  contains the snapshots of 1323 VMs.
%Since inner-VM deduplication is not involved in the first snapshot, this data set helps us to 
%study the CDS deduplication against user-related data. The overall size of dataset2 is 23.5 TB.
%
%using a dataset containing 10 snapshots of 35 VMs.
A VM image is  divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content blocks ~\cite{similar94} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
The final snapshots are stored in a distributed file system built on the same cluster and local I/O rate
is averaged about 50MB/second.  The seek cost of each random IO request is about  10 milliseconds.
%Popularity of data blocks are collected through global counting 
%and the top 1\% will fall into CDS, as discussed in Section~\ref{sect:crossVM}.

@@

Figure~\ref{fig:memory} shows the total parallel time in hours to backup 2500 VMs on
 a 100-node cluster a  when 
the memory limit imposed on each node varies  from 20MB to 200MB.
The overall processing time does not have a significant reduction as we increase to 200MB.
The parallel time includes the first scanning time, 
of dirty VM segments to generate block fingerprints, data redistribution for request accumulation,
fingerprint comparison, duplicate summary distribution, and the second scan time for real backup.
The total time is mainly dominated by  the two scans of dirty VM segments.
The time for copying to the final backup storage is overlapped with VM scanning.
%Each scan takes about 1.4 hours.
The aggregated throughput is  about 9.6GB per second,
which is the size of 2500 VM images divided by the parallel time. 
The result shows the backup with multi-stage deduplication  for all VM images can be 
completed in about 2.9 hours with the very small resource usage.
As we vary the cluster size and the number of VMs for backup remains a constant (25),
the processing time and throughput do not change much. Thus the aggregated throughput
scales up linearly.

Table~\ref{tab:overall} shows memory usage and performance when 35MB memory limit is imposed and
the number of parititons per machine ($q$) varies from 50 to 1500.
Row 2 shows the memory usage required to load a partition of global index and detection requests,
varying from 156MB to 5.2MB. With  35MB memory limit,  $q=50$ would not be a viable choice.
Row 3 is the total time and Row 4 is  the aggregated throughput of  100 nodes.
%The result shows the backup with multi-phase deduplication  for all VM images can be completed in about 3.09 hours
%with the very small resource usage.
Row 5 is  the parallel time if  we use Option 1 with $p\times q$ send buffers 
described in Section~\ref{sect:arch}. 
With the large number of buffers, the available space per buffer reduces significantly as $q$ increases, which leads to
a big increase of IO requests and seek cost.

\begin{table}[hbt]
\caption{ Performance when the number of partitions varies.}
\begin{center}
\begin{tabular} {|c|c|c|c|c|c|}
\hline \#Partitions  & 50 & 275  &450 &  750 &  1500 \\
\hline Index+request & 156&  28.4 & 17.4 & 10.7 & 5.2 \\
  (MB)           &  &   &  & &  \\

\hline Total Time  & N/A&  2.9 & 2.91 & 3 & 3.1 \\
  (Hours)           &  &   &  & &  \\
%\hline Throughput/node & N/A&  85MB/s& 90M & 87 & 81 \\
\hline Throughput GB/s& N/A&  9.6& 9.5 & 9.3 & 8.9 \\
\hline Total time & N/A&  8& 13.3 & 18.7 & 34 \\
  (Option 1)           &  &   &  & &  \\
\hline
\end{tabular}
\end{center}
\label{tab:overall}
\end{table}

>>>>

We have implemented and evaluated a prototype of our multi-stage deduplication scheme on a Linux cluster
of multi-core AMD Bulldozer FX8120 and Intel Nehalem E5530 machines.  
Our implementation is based on Alibaba's Xen cloud platform~\cite{Aliyun,WeiZhangIEEE}.

Objectives of our evaluation are:
1) Analyze the deduplication throughput and effectiveness for a large number of VMs.
%Compare with the data domain approach~\cite{bottleneck08}.
2) Examine the impacts of buffering during metadata exchange.

@@

%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
We have performed a trace-driven study using  a 1323 VM dataset  collected from 100 
Alibaba Aliyun's cloud nodes~\cite{WeiZhangIEEE}.
Each multi-core machine   hosts  up to 25 VMs. 
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.
The backup of VM snapshots is completed within a few  hours every night.
Based on our study of its production  data,  each VM has about  40GB of storage  data usage on average
including OS and user data disk.
Each VM image is  divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content blocks ~\cite{similar94} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
The final snapshots are stored in a distributed file system built on the same cluster and the average I/O bandwidth
of local storage is about 50MB/second.  The seek cost of each random IO request is about  10 milliseconds.
%and that translates to an aggregated backup throughput of 139GB per second, or 500TB per hour.
%  For 2 hours, every machine can do 139MB/second. For 3 hours, then 92MB/second

@@

%In this dataset, there are 10 snapshots per each VM user and the total amount of space 
%investigated is 17.5 terabytes.
%This dataset contas compose of 35 VMs from 7 popular OSes: 
%Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. For each OS, 
%5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
%The overall data size for this 700 full snapshots is 17.6 TB.
%This dataset  contains the snapshots of 1323 VMs.
%Since inner-VM deduplication is not involved in the first snapshot, this data set helps us to 
%study the CDS deduplication against user-related data. The overall size of dataset2 is 23.5 TB.
%
%using a dataset containing 10 snapshots of 35 VMs.
%Popularity of data blocks are collected through global counting 
%and the top 1\% will fall into CDS, as discussed in Section~\ref{sect:crossVM}.

@@

Figure~\ref{fig:memory} shows the total parallel time in hours to backup 2500 VMs on
 a 100-node cluster a  when 
the memory limit $M$ imposed on each node varies  from 20MB to 200MB.
The overall processing time does not have a significant reduction as $M$ increases to 200MB.
The parallel time includes the first scanning time, 
of dirty VM segments to generate block fingerprints, data redistribution for request accumulation,
fingerprint comparison, duplicate summary distribution, and the second scan time for real backup.
The total time is mainly dominated by  the two scans of dirty VM segments.
The time for copying to the final backup storage is overlapped with VM scanning.
%Each scan takes about 1.4 hours.
The aggregated throughput is  about 9.6GB per second,
which is the size of 2500 VM images divided by the parallel time. 
The result shows the backup with multi-stage deduplication  for all VM images can be 
completed in about 2.9 hours with 35MB memory and 8GB disk overhead.
As we vary the cluster size $p$,  the parallel time does  not change much, and  the aggregated throughput
scales up linearly since the number of VMs is  $25p$.

Table~\ref{tab:overall} shows performance change when memory limit $M$=35MB is imposed and
the number of parititons per machine ($q$) varies from 50 to 1500.
Row 2 is the memory space required to load a partition of global index and detection requests.
When $q=50$, the required memory is 156MB, which  is  not  a viable choice  under  constraint $M=$35MB.  
Row 3 is the parallel time and Row 4 is  the aggregated throughput of  100 nodes.
%The result shows the backup with multi-phase deduplication  for all VM images can be completed in about 3.09 hours
%with the very small resource usage.
Row 5 is  the parallel time if  we use Option 1 with $p\times q$ send buffers 
described in Section~\ref{sect:arch}. 
With the large number of buffers, the available space per buffer reduces significantly as $q$ increases, which leads to
a big increase of IO requests and seek cost.

\begin{table}[hbt]
\caption{ Performance when the number of partitions per machine node varies and $M$=35MB is imposed.}
\begin{center}
\begin{tabular} {|c|c|c|c|c|c|}
\hline \#Partitions  & 50 & 275  &450 &  750 &  1500 \\
\hline Index+request & 156&  28.4 & 17.4 & 10.7 & 5.2 \\
  (MB)           &  &   &  & &  \\

\hline Total Time  & N/A&  2.9 & 2.91 & 3 & 3.1 \\
  (Hours)           &  &   &  & &  \\
%\hline Throughput/node & N/A&  85MB/s& 90M & 87 & 81 \\
\hline Throughput GB/s& N/A&  9.6& 9.5 & 9.3 & 8.9 \\
\hline Total time & N/A&  7.8& 11.7 & 14.8 & 26 \\
  (Option 1)           &  &   &  & &  \\
\hline
\end{tabular}
\end{center}
\label{tab:overall}
\end{table}

### 2013-03-10 hotstorage13/intro.tex
revision

Among the three factors - time, cost and de-deduplication efficiency, one of them has to be compromised for the other two. For instance, if we were build a deduplication system that has a high rate of duplication detection and has a very fast response time, 
it would need a lot of memory to hold fingerprint index and do fast comparison.  This leads to a compromise on cost. 
The objective of our project is to lower the cost incurred while sustaining the highest de-duplication ratio
and  the sufficient deduplication  throughput for dealing with a large number of VM images. 
% This business model partially  resembles that of Amazon Glacier.

%Our solution  only uses a very small amount of memory space during backup while completing  snapshot backup and 
%deletion in a reasonable amount of time. 
The traditional approach to  deduplication for backup is an inline approach which follows
a sequence of VM block reading, duplicate detection,  and non-duplicate  block write to the 
backup storage.  
Our key idea  is to  first perform parallel duplicate detection for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs detection   partition by partition 
with minimal resource usage.
Fingerprint based partitioning allows highly parallel duplicate detection  and also simplifies 
reference counting management.  
% as the entire process can also be divided into a
%parition-wise  deletion. 
With careful parallelism and buffer  management,
this multi-phase detection scheme can provide  a sufficient throughput for VM backup.

>>>>

Among the three factors - time, cost and de-deduplication efficiency, one of them has to be compromised for the other two. For instance, if we were building a deduplication system that has a high rate of duplication detection and has a very fast response time, 
it would need a lot of memory to hold fingerprint index and cache.  This leads to a compromise on cost. 
The objective of our project is to lower the cost incurred while sustaining the highest de-duplication ratio
and  the sufficient deduplication  throughput for dealing with a large number of VM images. 
% This business model partially  resembles that of Amazon Glacier.

%Our solution  only uses a very small amount of memory space during backup while completing  snapshot backup and 
%deletion in a reasonable amount of time. 
The traditional approach to  deduplication for backup is an inline approach which follows
a sequence of block reading, duplicate detection,  and non-duplicate  block write to the 
backup storage.  
Our key idea  is to  first perform parallel duplicate detection for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs detection   partition by partition 
with minimal resource usage.
Fingerprint based partitioning allows highly parallel duplicate detection  and also simplifies 
reference counting management.  
% as the entire process can also be divided into a
%parition-wise  deletion. 
With careful parallelism and buffer  management,
this multi-stage detection scheme can provide  a sufficient throughput for VM backup.
